\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-07-29 syn-translate.tex
edits

The implementation described in \cref{sec:coqimpl} can be formalized in
our metatheory. The described compilation strategy can be considered as
a \textit{syntactic translation}, a function mapping from our syntactic
(QIIT-)model to a ``subpart''\footnote{Not exactly subpart---the only
difference lies in the "Wrec", in our metatheory, it is constructed
using linkage; in ``subpart'', it will be using sigma types (translation
of the linkage).} of our syntactic (QIIT-)model \textit{without
linkages}.\footnote{Here, syntactic QIIT-model can be roughly understood
as just AST, with equational rules acting like quotient. So this mapping
is mapping AST to AST, respecting quotient/judgemental equality.} This
``subpart'' is the standard MLTT with inductive type signature, a
(non-conventional) W-type and a recursor that using sigma type to
aggregate the handlers.

Basically, we can compile away the judgement $\goodSig{\Gamma}{v}{n}$,
the type $\cL$ and $\cC$, and related terms, by transforming them into
the corresponding sigma type.

@@

We use $\denotesT{}$ as the translation. This translation is invariant
under the subpart. Since we are in an intrinsic setting, everything here
is well-typed (well-formed)including this translation. Thus
$\denotesT{T}$ means the same as $\denotesT{\goodType{\Gamma}{T}{}}$
(i.e. every type is tracking its context). In other words, every
derivation of any judgement will track its ``context''. This will be
clearer in the (pseudo-)Agda formalization.

\begin{align*}
  \text{We will define a new type }& \goodType{\Gamma}{\Sigr{\Gamma}{n}}{} \text{ when } \goodCtx{\Gamma}{} \text{ and } n \in \mathbb{N} \\
  \denotesT{\goodSig{\Gamma}{\_}{n}},\ &({\Sigr{\Gamma}{n}}), \text{ and } \ \denotesT{\cC} \text{ are mutually recursively defined together, } \\
  & \text{and they define inductively on the signature length} \\  
  \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \iff \goodTerm{\Gamma}{\sigma}{\Sigr{\Gamma}{n}} \\ 
  {\Sigr{\Gamma}{(n+1)}} &= 
    \Sigma~(\Sigr{\Gamma}{n})
          ~(\Sigma~\cU
                  ~(\Sigma~(\Pi(\cC (\pi_2[\pi_1]))(El~\pi_2[\pi_1]))~\cU)) \\
  \Sigr{\Gamma}{0} &= \top \\
  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &\text{ defines upon } \denotesT{\sigma} \\
  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= 
    \Sigma~(\denotesT{\cC ("pjl"~\sigma)})~(El \ ("app"~("pjr"^3~\sigma))[(\pi_1, "app"~("pjl"~("pjr"^2~t)))]) \\
      &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\
  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= \top \quad
      \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}} \\ 
  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} & \text{ is defined upon } \denotesT{\sigma} \\
  & \text{ and inductively on the signature length} \\ 
  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} &=
  \denotesT{\cL("pjl"~\sigma)} \times \Pi(El~("pjl"~("pjr"~\sigma)))(El~("app"~("pjr"^3~\sigma))) \\
  &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\
  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} &= \top \quad \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}} \\
  \denotesT{\goodSig{\Gamma}{\nu^+~\sigma~f~T}{n+1}} &= (\sigma, "c"~A, "lam"~f, "lam"~("c"~T)) \\ 
  \denotesT{\goodTerm{\Gamma}{\mu^+~m~t}{\cL(\nu^+~\sigma~T)}} &= (m, "lam"~t) \\ 
  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC\sigma}} & \text{ is defined upon } \denotesT{m}\\
  & \text{ and inductively on the signature length} \\ 
  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC\sigma}} &= 
  (\denotesT{\cCt o}, t["sf"~f][("id", \denotesT{\cCt o})]) \\
  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = "pjl"~m, t = "app"~("pjr"~m), f = p_f\nu~\sigma \\ 
  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC \sigma}} &= () \quad \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}}
\end{align*}
\begin{align*}
  &\denotesT{\goodSig{\Gamma}{p_1\nu~\sigma}{n}} = "pjl"~\sigma 
  &&\denotesT{\goodType{\Gamma}{p_1\nu'~\sigma}{}} = El~("pjl"~("pjr"~\sigma)) \\
  &\denotesT{\goodSeal{\Gamma}{p_f\nu~\sigma}{p_1\nu~\sigma}{p_1\nu'~\sigma}} = "app"~("pjl"~("pjr"^2~\sigma)) 
  &&\denotesT{\goodSig{\Gamma}{p_2\nu~\sigma}{n}} = El~("app"~("pjr"^3~\sigma)) \\ 
  &\denotesT{\goodTerm{\Gamma}{p_1\mu~o}{\cL(p_1\nu\sigma)}} = "pjl"~o 
  &&\denotesT{\goodTerm{\Gamma, p_1\nu'~\sigma}{p_2\mu'~o}{p_2\nu~\sigma}} = "app"~("pjr"~o)
\end{align*}
\begin{align*}
  \denotesT{\goodInh{\Gamma}{\_}{\sigma_1}{\sigma_2}} &= \goodTerm{\Gamma, \cL\sigma_1}{\_}{\cL\sigma_2[\pi_1]} \\ \text{ and thus } \denotesT{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}} &\iff  \goodTerm{\Gamma, \cL\sigma_1}{h}{\cL\sigma_2[\pi_1]} \\
  \denotesT{\goodInh{\Gamma}{"inhinh"~h~T~\uparrow^s}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(\pi_1, \uparrow^s)])}} &= \mu^+~(h[(\pi_1,\pi_1\mu~\pi_2)])\\ &\quad \quad \{f_2[\pi_1]\}~(p_2\mu'~\pi_2)[(\pi_1, \uparrow^s[\pi_1^{\uparrow}])] \\
  \denotesT{\goodInh{\Gamma}{"inhov"~h~t}{(\nu^+ \  \sigma_1 \  T_1)}{(\nu^+ \  \sigma_2\  T_2)}} &= \mu^+ (h[(\pi_1, p_1\mu~\pi_2)])~\{f_2[\pi_1]\}~t[\pi_1^{\uparrow}]
\end{align*}

>>>>

The implementation described in \cref{sec:coqimpl} can be formalized in
our metatheory. The described compilation strategy can be considered as
a \textit{syntactic translation}, a function mapping from our syntactic
(QIIT-)model to a ``subpart''\footnote{Not exactly subpart---the only
difference lies in "Wrec". In our metatheory, it is constructed
using linkage; in ``subpart'', it will be using $\Sigma$-types (translation
of the linkage).} of our syntactic (QIIT-)model \textit{without
linkages}.\footnote{Here, syntactic QIIT-model can be roughly understood
as just AST, with equational rules acting like quotient. So this mapping
is mapping AST to AST, respecting quotient/judgemental equality.} This
``subpart'' is the standard MLTT with inductive type signatures, an
(unconventional) formulation of W-types,
and recursors that use $\Sigma$-types to aggregate handlers.

Basically, we can compile away the judgement $\goodSig{\Gamma}{v}{n}$,
the type $\cL$ and $\cC$, and related terms, by transforming them into
corresponding $\Sigma$-types.

@@

We use $\denotesT{}$ as the translation. This translation is invariant\YZ{What does "invariant" mean?}
under the subpart. Since we are in an intrinsic setting, everything here
is well-typed (well-formed)including this translation. Thus
$\denotesT{T}$ means the same as $\denotesT{\goodType{\Gamma}{T}{}}$
(i.e. every type is tracking its context). In other words, every
derivation of any judgement will track its ``context''. This will be
clearer in the (pseudo-)Agda formalization.

\begin{align*}
  \text{We will define a new type }& \goodType{\Gamma}{\Sigr{\Gamma}{n}}{} \text{ when } \goodCtx{\Gamma}{} \text{ and } n \in \mathbb{N} \\
  \denotesT{\goodSig{\Gamma}{\_}{n}},\ &({\Sigr{\Gamma}{n}}), \text{ and } \ \denotesT{\cC} \text{ are mutually recursive, } \\
  & \text{defined by induction on the signature length} \\  
  \denotesT{\goodSig{\Gamma}{\_}{n}} &= (\goodTerm{\Gamma}{\_}{\Sigr{\Gamma}{n}}) \text{, and thus } \denotesT{\goodSig{\Gamma}{\sigma}{n}} \iff \goodTerm{\Gamma}{\sigma}{\Sigr{\Gamma}{n}} \\ 
  {\Sigr{\Gamma}{(n+1)}} &= 
    \Sigma~(\Sigr{\Gamma}{n})
          ~(\Sigma~\cU
                  ~(\Sigma~(\Pi(\cC (\pi_2[\pi_1]))(El~\pi_2[\pi_1]))~\cU)) \\
  \Sigr{\Gamma}{0} &= \top \\
  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &\text{ defines upon } \denotesT{\sigma} \\
  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= 
    \Sigma~\denotesT{\cC ("pjl"~\sigma)}~(El \ ("app"~("pjr"^3~\sigma))[(\pi_1, "app"~("pjl"~("pjr"^2~t)))]) \\
      &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\
  \denotesT{\goodType{\Gamma}{\cC \sigma}{}} &= \top \quad
      \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}} \\ 
  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} & \text{ is defined upon } \denotesT{\sigma} \\
  & \text{ and inductively on the signature length} \\ 
  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} &=
  \denotesT{\cL("pjl"~\sigma)} \times \Pi(El~("pjl"~("pjr"~\sigma)))(El~("app"~("pjr"^3~\sigma))) \\
  &\text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}} \\
  \denotesT{\goodType{\Gamma}{\cL\sigma}{}} &= \top \quad \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}} \\
  \denotesT{\goodSig{\Gamma}{\nu^+~\sigma~f~T}{n+1}} &= (\sigma, "c"~A, "lam"~f, "lam"~("c"~T)) \\ 
  \denotesT{\goodTerm{\Gamma}{\mu^+~m~t}{\cL(\nu^+~\sigma~T)}} &= (m, "lam"~t) \\ 
  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC\sigma}} & \text{ is defined upon } \denotesT{m}\\
  & \text{ and inductively on the signature length} \\ 
  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC\sigma}} &= 
  (\denotesT{\cCt o}, t["sf"~f][("id", \denotesT{\cCt o})]) \\
  & \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{n+1}}, \text{ where } o = "pjl"~m, t = "app"~("pjr"~m), f = p_f\nu~\sigma \\ 
  \denotesT{\goodTerm{\Gamma}{\cCt m}{\cC \sigma}} &= () \quad \text{given } \denotesT{\goodSig{\Gamma}{\sigma}{0}}
\end{align*}
\begin{align*}
  &\denotesT{\goodSig{\Gamma}{p_1\nu~\sigma}{n}} = "pjl"~\sigma 
  &&\denotesT{\goodType{\Gamma}{p_1\nu'~\sigma}{}} = El~("pjl"~("pjr"~\sigma)) \\
  &\denotesT{\goodSeal{\Gamma}{p_f\nu~\sigma}{p_1\nu~\sigma}{p_1\nu'~\sigma}} = "app"~("pjl"~("pjr"^2~\sigma)) 
  &&\denotesT{\goodSig{\Gamma}{p_2\nu~\sigma}{n}} = El~("app"~("pjr"^3~\sigma)) \\ 
  &\denotesT{\goodTerm{\Gamma}{p_1\mu~o}{\cL(p_1\nu\sigma)}} = "pjl"~o 
  &&\denotesT{\goodTerm{\Gamma, p_1\nu'~\sigma}{p_2\mu'~o}{p_2\nu~\sigma}} = "app"~("pjr"~o)
\end{align*}
\begin{align*}
  \denotesT{\goodInh{\Gamma}{\_}{\sigma_1}{\sigma_2}} &= \goodTerm{\Gamma, \cL\sigma_1}{\_}{\cL\sigma_2[\pi_1]} \\ \text{ and thus } \denotesT{\goodInh{\Gamma}{h}{\sigma_1}{\sigma_2}} &\iff  \goodTerm{\Gamma, \cL\sigma_1}{h}{\cL\sigma_2[\pi_1]} \\
  \denotesT{\goodInh{\Gamma}{"inhinh"~h~T~\uparrow^s}{(\nu^+ \  \sigma_1 \  T)}{(\nu^+ \  \sigma_2 \  T[(\pi_1, \uparrow^s)])}} &= \mu^+~(h[(\pi_1,\pi_1\mu~\pi_2)])\\ &\quad \quad \{f_2[\pi_1]\}~(p_2\mu'~\pi_2)[(\pi_1, \uparrow^s[\pi_1^{\uparrow}])] \\
  \denotesT{\goodInh{\Gamma}{"inhov"~h~t}{(\nu^+ \  \sigma_1 \  T_1)}{(\nu^+ \  \sigma_2\  T_2)}} &= \mu^+ (h[(\pi_1, p_1\mu~\pi_2)])~\{f_2[\pi_1]\}~t[\pi_1^{\uparrow}]
\end{align*}
